Quality Criteria for SKOS Vocabularies

### 2012-03-20 criteria.tex
reworked model definition

\begin{definition}[SKOS Vocabulary] Let the fully entailed RDFS interpretation~\cite{RDFSEM2012} (enriched by entailment of OWL inverse properties) of a SKOS vocabulary be a tuple of the form $V = \langle IR, C, SR, LV \rangle$, with

@@

\begin{itemize}
	\item \(IR = I_{CEXT}(\texttt{rdfs:Resource}^\mathcal{I})\) be the set of all resources in V, as defined in \cite{RDFSEM2012}.

	\item \(C = I_{CEXT}(\texttt{skos:Concept}^\mathcal{I})\) be the set of \textbf{conceptual resources}.

@@

	\item \(SR = I_{EXT}(\texttt{skos:semanticRelation}^\mathcal{I})\) be the set of \textbf{semantic relations} associating conceptual resources with one another.

	\item $LV = I_{CEXT}(\texttt{rdfs:Literal}^\mathcal{I})$

@@

The quality criteria we identified have in common that they are applied on a given SKOS vocabulary and identify a set of resources, which do not fulfill that criterion. Therefore, we can regard a quality criterion as being an abstract filtering function $F$, mapping a given vocabulary $V$ to the powerset $\mathfrak{P}(IR)$ of its resources.\todo{CM}{think about it and try to formalize this.} Each quality criterion we found can be implemented as a realization of such an abstract filtering function. Since the return value is a set of resources, statistical analyses can be computed on the number of resources in that set.

@@

\subsection{Labeling and Documentation Criteria}
The criteria defined in this section..\todo{CM}{describe}. Their filtering function \(f:V \rightarrow \mathfrak{P}(IR)\) yields all sets of \textbf{conceptual resources} violating the respective criterion.

@@

\subsubsection{Cyclic Hierarchical Relations} have been mentioned numerous times in thesaurus development literature. \cite{Soergel2002} suggests a ``check for hierarchy cycles" since they ``throw the program for a loop in the generation of a complete hierarchical structure''. However, SKOS does not define a formal constraint regarding hierarchy cycles but the SKOS reference document mentions them as potential problems. The SKOS documentation also does not state, how exactly hierarchical relationships are to be interpreted. There exist common forms like, e.g., ``generic-specific'', ``instance-of'' or ``whole-part'' \cite{Hedden2010,Harpring2010,Aitchison2000} where cycles would be considered a logical contradiction. To calculate this criterion, we assume all OWL inverse properties of the SKOS ontology properly entailed and construct two graphs, $G_{br}$ and $G_{nar}$, with the set of nodes being C. The set of edges in $G_{br}$ are all pairs identified by $I_{EXT}(\texttt{skos:broaderTransitive}^\mathcal{I})$, likewise those of $G_{nar}$ are $I_{EXT}(\texttt{skos:narrowerTransitive}^\mathcal{I})$. For each graph we now identify those nodes that are part of a cycle and return the set of strongly connected components they are contained in.

@@

\subsubsection{Undefined SKOS Property Usage}\todo{CM}{Merge 15 and 16}

>>>>

\begin{definition}[SKOS Vocabulary] Let the fully entailed RDFS interpretation~\cite{RDFSEM2012} (enriched by entailment of \texttt{owl:inverseOf} properties) of a SKOS vocabulary be a tuple of the form $V = \langle IR, C, SR, LV \rangle$, with

@@

\begin{itemize}
	\item \(IR = I_{CEXT}(\texttt{rdfs:Resource}^\mathcal{I})\) being the set of resources,

	\item $LV = I_{CEXT}(\texttt{rdfs:Literal}^\mathcal{I})$ being the set of literal values,

	\item \(C = I_{CEXT}(\texttt{skos:Concept}^\mathcal{I})\) being the set of \textbf{conceptual resources},

@@

	\item \(SR = I_{EXT}(\texttt{skos:semanticRelation}^\mathcal{I})\) being the set of \textbf{semantic relations} associating conceptual resources with one another.

@@

The quality criteria we identified have in common that they are applied on a given SKOS vocabulary and identify a set of resources, which do not fulfill that criterion. Therefore, we can regard a quality criterion as being an abstract filtering function $f:V \rightarrow \mathfrak{P}(IR)$, mapping a given vocabulary to the powerset of its resources. Each quality criterion we found can be implemented as a realization of such an abstract filtering function. Since the return value is a set of resources, statistical analyses can be computed on the number of resources in that set.

@@

\subsection{Labeling and Documentation Criteria}
The criteria defined in this section primarily focus on supporting interaction with human users that either work with the vocabulary itself or interact with a system using the vocabulary for, e.g., search and retrieval purposes.

@@

\subsubsection{Cyclic Hierarchical Relations} have been mentioned numerous times in thesaurus development literature. \cite{Soergel2002} suggests a ``check for hierarchy cycles" since they ``throw the program for a loop in the generation of a complete hierarchical structure''. However, SKOS does not define a formal constraint regarding hierarchy cycles but the SKOS reference document mentions them as potential problems. The SKOS documentation also does not state, how exactly hierarchical relationships are to be interpreted. There exist common forms like, e.g., ``generic-specific'', ``instance-of'' or ``whole-part'' \cite{Hedden2010,Harpring2010,Aitchison2000} where cycles would be considered a logical contradiction. To calculate this criterion, we construct two graphs, $G_{br}$ and $G_{nar}$, with the set of nodes being C. The set of edges in $G_{br}$ are all pairs identified by $I_{EXT}(\texttt{skos:broaderTransitive}^\mathcal{I})$, likewise those of $G_{nar}$ are $I_{EXT}(\texttt{skos:narrowerTransitive}^\mathcal{I})$. For each graph we now identify those nodes that are part of a cycle and return the set of strongly connected components they are contained in.

@@

\subsubsection{Undefined SKOS Property Usage}
Some vocabularies reference resources in the SKOS namespace that cannot be resolved. A common reason for this is, that there seems to exist a misunderstanding about the correct way of introducing new classes or properties in an OWL document. In some cases, vocabularies ``invent'' new terms in the SKOS namespace in order to meet certain requirements of the publishing organizations that cannot (yet) be expressed within the SKOS schema. Another reason for the presence of unresolvable resource references in the SKOS namespace is that the vocabulary might be outdated. With the SKOS W3C recommendation maturing, some properties have been removed from the current version of the schema. Vocabularies using these ``deprecated'' properties hence contain parts of their information in a non ``standard"-compliant way, having negative effects when, e.g., crawling vocabularies to gather specific information.
